Ablation study on the residual adapter in Neural Machine Translation

### 2020-08-14 2020/wmt_20/emnlp2020.tex
developping

\subsection{Architecture \label{ssec:architecture}}
\fyTodo{More contexts and notations from the transformer}

Residual adapters modify context vectors in each layer by applying several transformations as follows:\fyTodo{Use align env}
\begin{align*}
  h^{i}_1 &= \mathbf{W}_{d_{model}}^{d_{bottleneck}}h^{i} + b_{1} \\
  h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
  h^{i}_3 &= \mathbf{W}_{d_{bottleneck}}^{d_{model}}h_2^{i} + b_{2} \\
  \bar{h}^{i} &= h^{i}_3 + h^i
\end{align*}
For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}\fyTodo{attention aux matrices $W_i$}

The "adapted" context vector $\bar{h}^i$ is the input of the $(i+1)^{\text{th}}$ attention\fyDone{Self attention ?} layer. In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.

@@

\subsection{Highway ...}
Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers. We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}. The benefit of this variant is that we can manipulate all values of the adapters at once. An disadvantage of this variant is a lower complexity\fyTodo{Capicity ?? justify this ?} than original version because the change of context vector only occurs in the end of the stack of layers.

@@

Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De. In most cases, the performance increases with respect to the number of residual adapters used in architecture. The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level. However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?} 
\begin{table*}
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
    \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 \\
   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}\fyTodo{Boldface ?}
  \label{tab:performance-en-fr}
\end{table*}

@@

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ \hline 
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
    \system{MDL-HW-res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7\\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems.}
  \label{tab:performance-multi}
\end{table*}

@@

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} \\ \hline  
    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\
   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 \\ \hline 
   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 27.13\\ 
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1 & 35.92   \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9 & 34.91 \\
    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          & 37.01 \\
    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} & 31.78 \\
    \system{wada}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems}
  \label{tab:performance-random}
\end{table*}

\section{Related Work \label{sec:related}}
\mpTodo{related work}
\section{Discussion \label{sec:discussion}}
\mpTodo{discussion}
\section*{Acknowledgments}

\bibliographystyle{acl_natbib}
\bibliography{emnlp2020}
\appendix
\section{Appendices}
\label{sec:appendix}
\mpTodo{appendix}
\section{Supplemental Material}
\label{sec:supplemental}

>>>>

\subsection{Architecture \label{ssec:architecture}}
\fyTodo{More contexts and notations from the transformer}
For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension d. The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyTodo{Use align env}
\begin{align*}
  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
  h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
  \bar{h}^{i} &= h^{i}_3 + h^i
\end{align*}
Where matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension. For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$.\fyDone{or is it $h_i$ ?}\fyTodo{attention aux matrices $W_i$}

The "adapted" context vector $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is number of layer  is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side. In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder \footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter}.

@@

\subsection{Highway ...}
Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers. We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}. The benefit of this variant is that we can manipulate all values of the adapters at once.

@@

Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De. In most cases, the performance increases with respect to the number of residual adapters used in architecture. The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level. However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|l|l|} \hline
    Model & params \\ \hline 
    \system{Transformer}  & 65 \\
   \system{Residual Adapter} & 1 \\
     \hline
  \end{tabular}
  \caption{Number of parameters($\times 10^3$)}
  \label{tab:params}
\end{table*}

\begin{table*}
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
    \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 \\
   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}\fyTodo{Boldface ?}
  \label{tab:performance-en-fr}
\end{table*}

@@

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\% & 23.5\\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} \\
   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} \\ \hline 
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1    \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9   \\
    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          \\
    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} & 49.7   \\
    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 43.4 \\ 
    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 47.5 \\
    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} \\
    \system{MDL-HW-res}   & 37.38&	56.43&	52.13&	33.7&	44.81&	89.77&	52.4 \\ 
    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 33.7 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems.}
  \label{tab:performance-multi}
\end{table*}

@@

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{RND}} \\ \hline  
    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 & 49.4 \\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} &  \SB{53.8} & 32.55 \\
   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  &  \SB{53.3} & 38.38 \\ \hline 
   \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 & 27.13\\ 
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5}  & 50.1 & 35.92   \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} & 49.9 & 34.91 \\
    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & 49.8          & 37.01 \\
    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & \SB{51.9} & 31.78 \\
    \system{wada}    & 37.98 &	57.46&	53.05&	33.55&	46.01&	90.07&	53.0&	48.96 \\
    \system{FT-full-UEW} &	37.94 &	55.97 &	52.1 &	33.71 &	44.91 &	89.08 &	52.3 &	46.48 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems}
  \label{tab:performance-random}
\end{table*}

\section{Related Work \label{sec:related}}

\section{Discussion \label{sec:discussion}}
\mpTodo{discussion}
%\section*{Acknowledgments}

\bibliographystyle{acl_natbib}
\bibliography{emnlp2020}
%\appendix
%\section{Appendices}
%\label{sec:appendix}
%\mpTodo{appendix}
%\section{Supplemental Material}
%\label{sec:supplemental}
